---
title: "Correlation between TCB & TSB"
author: "Matthew Moreno"
date: "5/24/2019"
output: 
    html_document:
        toc_depth: 2
        toc: TRUE
        toc_float: TRUE 
        smooth_scroll: TRUE
        theme: yeti
        df_print: paged
---

```{r , echo = FALSE , message = FALSE }

library( dplyr )

library( magrittr )

library( ggplot2 )

library( ggpubr)

library( tibble )

library( readr )

library( pander )

library( qqplotr )

library( lmtest )

library( boot )

library( Hmisc )

library( gridExtra)

library( forcats )

library( corrplot)

library( BSDA )

library( DT )



setwd("~/Desktop/Stats/183/Projecto")



Nursery_Data <- read_csv( file = "Nursery- Jan, Feb, March 2014.csv" )  %>%  as_tibble()  %>% na.exclude()


Cor_Data <- read_csv( file = "Correlation between TSB and TCB Apri May and June 2014.csv")  %>% as_tibble()  %>% na.exclude()


Nursery_Data$Gender <- Nursery_Data$Gender  %>% as.factor() 
Cor_Data$Gender     <- Cor_Data$Gender      %>% as.factor()


Cor_Data$`MR No`     <- Cor_Data$`MR No`      %>% as.character()


levels( Nursery_Data$Gender )[ levels( Nursery_Data$Gender ) == "f" ] <- "F"
levels( Nursery_Data$Gender )[ levels( Nursery_Data$Gender ) == "m" ] <- "M"
levels( Cor_Data$Gender )[ levels( Cor_Data$Gender ) == "f" ] <- "F"
levels( Cor_Data$Gender )[ levels( Cor_Data$Gender ) == "m" ] <- "M"




T_Data <- Nursery_Data  %>% 
    full_join( Cor_Data)  %>% 
    select( TCB , TSB , `Birth Wt` , Gender )

attach( T_Data)

```

&nbsp;

# *Exploratory Data Analysis*

&nbsp;

## *Density Distributions of TSB & TCB*

```{r , echo = FALSE}


.TSB_Max_Density <- density(T_Data$TSB)$x[143]   %>%  round( digits = 2)



.TCB_Max_Density <- density(T_Data$TCB)$x[125]  %>% round( digits = 2)


ggplot( data = T_Data ) +
    geom_density( aes( x = TSB ) , fill = "#90D6D7" , alpha = 0.5) +
    geom_density( aes( x = TCB) , fill = "#EFB1AF" , alpha = 0.5 ) +
    geom_vline( xintercept = .TSB_Max_Density , lty = 3 ) +
    geom_vline( xintercept = .TCB_Max_Density , lty = 3 ) +
    labs( x = "Bilirubin Levels" , y = "Density") +
    annotate( geom = "text" , label = "TSB" , x = 5.8 , y = 0.45  , color = "blue") +
    annotate( geom = "text" , label = "TCB" , x = 9  , y = 0.45 , color = "red") +
    annotate( geom = "text" , x = 5.8 , y = 0.40 , label = .TSB_Max_Density , color = "blue") +
    annotate( geom = "text" , x = 9 , y = 0.40 , label = .TCB_Max_Density , color = "red") +
    ggsave("TSB TCB Density.png")






```

&nbsp;

## *Density Distributions of TSB & TCB by Gender*

&nbsp;

```{r , echo = FALSE}

ggplot( data = T_Data , aes( color = Gender) ) +
    geom_density( aes( x = TSB , fill = Gender) , alpha = 0.2) +
    geom_density( aes( x = TCB , fill = Gender) , alpha = 0.2 ) +
    labs( x = "Bilirubin Levels" , y = "Density") +
    annotate( geom = "text" , label = "TSB" , x = 6.65 , y = 0.5 ) +
    annotate( geom = "text" , label = "TCB" , x = 8.4  , y = 0.45 ) +
    theme( legend.position = "bottom") 
    ggsave( "TSB TCB Gender Density.png")



```

&nbsp;

# *Classical Statistical Tests*

&nbsp;

## *Two Sample T-Test Based on Gender*

&nbsp;

```{r , echo = FALSE}


.Female <- T_Data  %>% 
    filter( Gender == "F" )



.Male <- T_Data  %>% 
    filter( Gender == "M" )








t.test( x = .Male$TCB , y = .Female$TCB )  %>% pander()

t.test( x = .Male$TSB , y = .Female$TSB )  %>% pander()



```




&nbsp;


```{r , echo = FALSE}





.TSB_Box <- ggplot( data = T_Data , aes( fill = Gender )) +
    geom_boxplot( aes(  y = TSB) , alpha = .5 , show.legend = F ) 


.TCB_Box <- ggplot( data = T_Data , aes( fill = Gender )) +
    geom_boxplot( aes(  y = TCB) , alpha = 0.5  ) +
    theme( legend.position = "bottom" )




.Legend <- get_legend(.TCB_Box)

.TCB_Box <- .TCB_Box + theme( legend.position = "none" )


grid.arrange( .TSB_Box , .TCB_Box , .Legend , nrow = 2 , ncol = 2 , 
              layout_matrix = rbind(c(1,2), c(3,3)),
             widths = c(2.7, 2.7), heights = c(2.5, 0.2))





```

&nbsp;

## *Two Sample Paired T-Test: Difference of means.*

&nbsp;

```{r , echo = FALSE}

t.test( x = TSB , y = TCB , mu = 0 , paired = TRUE )  %>% pander()

```


&nbsp;

## *Kolmogorov Smirnov Test*

&nbsp;

```{r , echo = FALSE , message = FALSE , warning = FALSE}
ks.test( x = TSB , y = TCB  )  %>% pander()
```





&nbsp;

# *TCB & TSB Plot*

&nbsp;


```{r , echo = FALSE , message = FALSE}


ggplot( data = T_Data , aes( x = TCB , y = TSB , color = Gender )) +
    geom_point() +
    geom_smooth( se = FALSE , method = "lm") +
    stat_regline_equation( size = 5 ) +
    stat_cor( label.x.npc = 0.35 , label.y.npc = .95 , size = 4 ) +
    labs( x = "TCB" , y = "TSB" , title = "TCB & TSB Relationship between Gender." ) +
    theme( plot.title = element_text( hjust = 0.5 )) +
    ggsave("TCB TSB Gender LM.eps")
    


```

&nbsp;

# *Correlation*

&nbsp;

## *TSB & TCB Correlation*

&nbsp;

```{r , echo = FALSE}

cor.test( TSB, TCB )  %>% pander()

```

&nbsp;

## *TSB & Birth Weight Correlation*

&nbsp;


```{r , echo = FALSE}

cor.test( TSB , `Birth Wt`)  %>% pander()

```

&nbsp;

## *TCB & Birth Weight Correlation*

&nbsp;


```{r , echo = FALSE}

cor.test( TCB , `Birth Wt`)  %>% pander()

```

&nbsp;

## *TCB & Gender Correlation*

&nbsp;

```{r , echo = FALSE , message = FALSE}


.Corr <- cor( x =  as.matrix(T_Data[1:3]))

p.mat <- cor.mtest(T_Data[1:3])$p


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))


corrplot( .Corr, method = "color" ,  col = col(200) , type="upper" , order = "hclust",  addCoef.col = "black", tl.col = "black", tl.srt = 45,  p.mat = p.mat , sig.level = 0.01 , insig = "blank", diag = T )

```

&nbsp;

# *Model Building*

&nbsp;

## *Simple Linear Model 1: TSB ~ TCB*

&nbsp;

```{r , echo = FALSE}

Simple_GLM_1 <- glm( TSB ~ TCB , data = T_Data , family = gaussian())

summary( Simple_GLM_1 )

```

&nbsp;

## *Simple Linear Model 2: TSB ~ TCB + Birth WT*

&nbsp;

```{r , echo = FALSE}

Simple_GLM_2 <- glm( TSB ~ TCB + `Birth Wt` , data = T_Data , family = gaussian())

summary( Simple_GLM_2 )

```


&nbsp;

## *Simple Linear Model 3: TSB ~ TCB + Birth WT + Gender*

&nbsp;

```{r , echo = FALSE}

Simple_GLM_3 <- glm( TSB ~ TCB + `Birth Wt` + Gender , data = T_Data , family = gaussian())

summary( Simple_GLM_3 )


```

&nbsp;

## _Simple Linear Model 4: TSB ~ TCB * Gender_

&nbsp;

```{r , echo = FALSE}

Simple_GLM_4 <- glm( TSB ~  Gender * TCB, data = T_Data , family = gaussian())

summary( Simple_GLM_4 )

```




&nbsp;

## _Simple Linear Model 5: TSB ~ TCB * Birth Wt_

&nbsp;


```{r , echo = FALSE}

Simple_GLM_5 <- glm( TSB ~  `Birth Wt` * TCB, data = T_Data , family = gaussian())

summary( Simple_GLM_5 )

```


# Model Conclusions

&nbsp;

__The Best preforming model to predict TSB levels in neonates is to only use TCB, since gender and birth weight are insignificant predictors in the model.__

&nbsp;


# *Model 1 Assumptions Check:*

&nbsp;

## *Normality of Residuals*

&nbsp;


```{r , echo = FALSE}

ggplot( mapping =  aes( sample = Simple_GLM_1$residuals )) +
    stat_qq_point() +
    stat_qq_line( color = "red" ) + 
    stat_qq_band( alpha = 0.35 ) +
    labs( x = "Theoretical" , y = "Sample" , title = "Q-Q Plot for Normality" ) +
    theme( plot.title = element_text( hjust = 0.5 )) +
    ggsave( "M1 QQ.png")



```

&nbsp;

```{r , echo = FALSE}

shapiro.test( resid( Simple_GLM_1))  %>% pander()

```


&nbsp;

## *Equal Variance*

&nbsp;

```{r , echo = FALSE}

ggplot( mapping =  aes( x = Simple_GLM_1$fitted.values , y = Simple_GLM_1$residuals )) +
    geom_point() +
    geom_hline( yintercept = 0 , color = "red" , lwd = 0.80 ) +
    labs( x = "Fitted Values" , y = "Residuals" , title = "Residuals VS Fitted Values") +
    theme( plot.title = element_text( hjust = 0.5 )) +
    ggsave( "M1 RF.eps")


```

&nbsp;

```{r , echo = FALSE }
bptest( Simple_GLM_1 )  %>% pander()
```


&nbsp;

## *Independent Residuals.*

&nbsp;


```{r , echo = FALSE}
dwtest( Simple_GLM_1 )  %>% pander()
```

&nbsp;

# *10 Fold Cross Validation*

&nbsp;

```{r , echo = FALSE }

cv.glm( data = T_Data , glmfit = Simple_GLM_1 , K = 10)$delta 

```

&nbsp;

# *GLM With Log Transformation*

&nbsp;

```{r}

Simple_GLM_1_T <- glm( log(TSB) ~ TCB , data = T_Data , family = gaussian())

summary( Simple_GLM_1_T )


```



# *Model 1 Transformed Assumptions Check:*

&nbsp;

## *Normality of Residuals*

&nbsp;


```{r , echo = FALSE}

ggplot( mapping =  aes( sample = Simple_GLM_1_T$residuals )) +
    stat_qq_point() +
    stat_qq_line( color = "red" ) + 
    stat_qq_band( alpha = 0.35 ) +
    labs( x = "Theoretical" , y = "Sample" , title = "Q-Q Plot for Normality" ) +
    theme( plot.title = element_text( hjust = 0.5 )) +
    ggsave( "M2 QQ.png")



```

&nbsp;

```{r , echo = FALSE}

shapiro.test( resid( Simple_GLM_1_T))  %>% pander()

```

&nbsp;

## *Equal Variance*

&nbsp;

```{r , echo = FALSE}

ggplot( mapping =  aes( x = Simple_GLM_1_T$fitted.values , y = Simple_GLM_1_T$residuals )) +
    geom_point() +
    geom_hline( yintercept = 0 , color = "red" , lwd = 0.80 ) +
    labs( x = "Fitted Values" , y = "Residuals" , title = "Residuals VS Fitted Values") +
    theme( plot.title = element_text( hjust = 0.5 )) +
    ggsave( "M2 RF.eps")


```

&nbsp;


```{r , echo = FALSE }
bptest( Simple_GLM_1_T )  %>% pander()
```

&nbsp;


## *Independent Residuals*

&nbsp;


```{r , echo = FALSE}
dwtest( Simple_GLM_1_T )  %>% pander()
```

&nbsp;

# *10 Fold Cross Validation*

&nbsp;

```{r , echo = FALSE }

cv.glm( data = T_Data , glmfit = Simple_GLM_1_T , K = 10)$delta 

```



&nbsp;



# *TSB 25th & 75th Quantile*

&nbsp;

### *TSB*

&nbsp;

```{r , echo = FALSE}

.TSB_Summary <- summary( TSB )

.TSB_Summary %>% pander()


```

&nbsp;

### *TCB*

&nbsp;

```{r , echo = FALSE}

.TCB_Summary <- summary( TCB )

.TCB_Summary %>% pander()


```
&nbsp;

## *Sensitivity*

&nbsp;

**The sensitivity is $0.9578$**

&nbsp;

```{r , echo = FALSE}

TCB_75th <- T_Data  %>% 
    filter( TCB < quantile( TCB , 0.75))


TSB_95th <- TCB_75th  %>% 
    filter( TSB > quantile(TSB , 0.95))

Sensitivity <-  (nrow(TCB_75th) / (nrow(TCB_75th) + nrow(TSB_95th)))


```

&nbsp;

## *Specificity*

&nbsp;

**The specificity is $0.5172$**

&nbsp;

```{r , echo = FALSE}

TCB_75th <- T_Data  %>% 
    filter( TCB > quantile( TCB , 0.75))


TSB_95th <- TCB_75th  %>% 
    filter( TSB < quantile(TSB , 0.95))


Specificity <-  (nrow(TCB_75th) / (nrow(TCB_75th) + nrow(TSB_95th)))

```

&nbsp;

## *Confusion Matrix*

&nbsp;

|   | + | - |
|---|---|---|
| + | 91  | 28  |
| - | 4  | 30  |

&nbsp;


# *Model Comparisons*

&nbsp;

## *Model 1 & Model 3*

&nbsp;


```{r}
anova( Simple_GLM_1 , Simple_GLM_3 , test = "LRT")
```

&nbsp;

## *Model 1 Transformed & Model 3*

&nbsp;

```{r}
Simple_GLM_3 <- glm( log(TSB) ~ TCB + `Birth Wt` + Gender , data = T_Data , family = gaussian())


anova( Simple_GLM_1_T , Simple_GLM_3 , test = "LRT") 



```


&nbsp;

# *Appendix*

&nbsp;

```{r}
T_Data  %>% DT::datatable()
```





